Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-03 paper/disc20clustering.tex
Restructure intro

\noindent \textbf{Motivation and model.}
The \emph{balanced repartitioning} problem (\OBRP{})
is a fundamental learning problem
that finds applications in the context of
distributed systems optimization~\cite{repartition-disc}. We are given a set $V$ of $n$ nodes 
(e.g., virtual machines or processes),
initially arbitrarily partitioned into $\ell$~clusters
(e.g., servers or entire racks),
each of size~$k$.
The nodes interact using
a sequence of pairwise communication requests
$\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a~certain amount of data.
Nodes in $C \subset V$ are \emph{collocated}
if they reside in the same cluster.

@@

In addition to the general model, we study a simplified Perfect Partition model (also studied by Henzinger et al.~\cite{sigmetrics19_partitioning} (SIGMETRICS 2019)) that we consider for a stronger lower bound.
There, we assume that $\sigma$
reveals the edges of a static graph
that can be perfectly partitioned
(no external requests are required).

\noindent \textbf{Contributions.}
We provide an improved lower bound 
of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
even in the learning model;
the best known lower bound so far was $\Omega(k)$,
even for a more general model~\cite{repartition-disc}.
We also present an asymptotically optimal, 
$O(k\cdot \ell)$-competitive algorithm
for the learning model.
For the repartitioning model, we present  
an asymptotically optimal,
$\Theta(\ell)$ algorithm for $k=3$;
the best known upper bound 
so far was $O(\ell^2)$~\cite{repartition-disc}.
Then, we present a strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\ell^2)$ additive constant..
In addition, we show a new, improved analysis of the algorithm from \cite{repartition-disc} that depends on the solutions to an interesting combinatorial problem (see Section~\ref{ssec:cascade}, The Cascade Problem).
%
The overview of known results is in Table~\ref{tab:overview}.

\section{Related work}

@@

@@

@@

Across this paper, we often refer to groups of communicating nodes.
We use this concept slightly differently in the lower bound and the upper bounds.
In our algorithms, we group communicating nodes into \emph{components}, and in the lower bound, we group nodes that may communicate into \emph{ground sets}.

@@

\noindent
\textbf{Ground sets.}
In the lower bound, a~\emph{ground set} is a group of nodes that start communicating if an algorithm splits them (places at different clusters).
Their role is to ensure that the algorithm maintains a partition of ground sets.
We emphasize that ground sets are revealed only upon their nodes are split.
\maciek{Other names: division, segmentation, alignment, formation}

@@

In this section we consider a variant of Online Balanced Partitioning, where requests are permanent: once a pair of nodes start communicating, they continue forever.
Each communicating pair must be collocated, as otherwise the cost would be infinite.
We study the instances with finite cost: where the communication graph can be perfectly partitioned in the clusters.

>>>>

\maciek{Practical motivation}

\subsection{Model.}

\maciek{We study two models. One is a subproblem.}

\noindent
\textbf{Partitioning model.}
The \emph{balanced repartitioning} problem (\OBRP{})
is a fundamental learning problem
that finds applications in the context of
distributed systems optimization~\cite{repartition-disc}. We are given a set $V$ of $n$ nodes 
(e.g., virtual machines or processes),
initially arbitrarily partitioned into $\ell$~clusters
(e.g., servers or entire racks),
each of size~$k$.
The nodes interact using
a sequence of pairwise communication requests
$\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a~certain amount of data.
Nodes in $C \subset V$ are \emph{collocated}
if they reside in the same cluster.

@@

\noindent
\textbf{Learning model.}
In addition to the general model, we study a simplified Perfect Partition model (also studied by Henzinger et al.~\cite{sigmetrics19_partitioning} (SIGMETRICS 2019)) that we consider for a stronger lower bound.
There, we assume that $\sigma$
reveals the edges of a static graph
that can be perfectly partitioned
(no external requests are required).

\subsection{Related work}

@@

\maciek{Offline results}
\maciek{Links to paging}

@@

\subsection{Our Contributions.}
We provide an improved lower bound 
of $\Omega(k\cdot\ell)$ on the competitive ratio of any online deterministic online algorithm 
even in the learning model;
the best known lower bound so far was $\Omega(k)$,
even for a more general model~\cite{repartition-disc}.
We also present an asymptotically optimal, 
$O(k\cdot \ell)$-competitive algorithm
for the learning model.
For the repartitioning model, we present  
an asymptotically optimal,
$\Theta(\ell)$ algorithm for $k=3$;
the best known upper bound 
so far was $O(\ell^2)$~\cite{repartition-disc}.
Then, we present a strictly $6$-competitive algorithm for $k=2$ that improves upon previous $7$-competitive algorithm with $O(\ell^2)$ additive constant..
In addition, we show a new, improved analysis of the algorithm from \cite{repartition-disc} that depends on the solutions to an interesting combinatorial problem (see Section~\ref{ssec:cascade}, The Cascade Problem).
%
The overview of known results is in Table~\ref{tab:overview}.

@@

Across this paper, we often refer to groups of communicating nodes.
We use this concept slightly differently in the lower bound and the upper bounds.
In our algorithms, we group nodes with a communication history into \emph{components}, and in the lower bound, we group nodes that may communicate into \emph{ground sets}.

@@

\noindent
\textbf{Ground sets.}
In the lower bound, a~\emph{ground set} is a group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
Their role is to ensure that the algorithm maintains a partition of ground sets into clusters.
We emphasize that ground sets are revealed only upon their nodes are split.
\maciek{Other names: division, segmentation, alignment, formation}

@@

In this section we consider a variant of Online Balanced Partitioning, where communication requests are permanent: once a pair of nodes start communicating, they continue forever.
To avoid infinite cost, each communicating pair must be collocated.
We study the instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters.
The algorithm's goal is to \emph{learn} the communication graph's partition.
